---
title: "shiny output"
author: "Yueran Cao"
date: "2025-03-08"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(install.packages.check.source = "no")
options(repos = c(CRAN = "https://cloud.r-project.org"))

```


```{r, results = "hide"}
##run plot
#original dataset
library(dplyr)
library(ggplot2)
library(RColorBrewer)
feature_table <- readRDS("data/data17")
cols_meta <- 2
df <- feature_table[, -c(1:cols_meta)]
df_ft<- cbind(feature_table[, 1:cols_meta] , df)
metadata <- feature_table[, 1:cols_meta]


#randomly assign sample types 
sample_types <- c("S1", "S2", "S3", "S4")
metadata$Sample_type <- sample(sample_types, nrow(metadata), replace = TRUE)


# 
install.packages("missMDA")
library(missMDA)

# PCA analysis
imputed_data <- imputePCA(df, ncp = 2)  
df_imputed <- imputed_data$completeObs 

pca_result <- prcomp(df_imputed, scale = TRUE)
pca_nc <- data.frame(pca_result$x)  
pca_nc <- pca_nc %>%
  mutate(sum_sample_int = rowSums(df_imputed))


pca_nc <- cbind(pca_nc, metadata)  # combine batch information

# ANOVA analysis
pc1_anova_nc <- summary(aov(PC1 ~ batch, data = pca_nc))
pc2_anova_nc <- summary(aov(PC2 ~ batch, data = pca_nc))
int_anova_nc <- summary(aov(sum_sample_int ~ batch, data = pca_nc))


# Determine the number of unique batches 
batch_num <- length(unique(metadata$batch))

# Set up color indices for the plots 
if(batch_num < 9) { 
  color_index <- brewer.pal(8, "Dark2")[1:batch_num] 
  } else { 
    color_index <- rep(brewer.pal(8, "Dark2"), 100)[1:batch_num] } 

# Print the color indices 
print(color_index)

# Ensure necessary libraries are loaded
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
library(ggplot2)

# Initialize plot list
p <- list()

# Plot 1: Sum Intensity for All Samples

p[[1]] <- ggplot(pca_nc, aes(y = sum_sample_int, x = 1:length(sum_sample_int), fill = as.factor(batch))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = color_index) +
  labs(x = "Sample number", y = "Sum Intensity", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("Sum intensity for all samples; p=", round(int_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 2: PC1 vs PC2 Colored by Batch
p[[2]] <- ggplot(pca_nc, aes(x = PC1, y = PC2, color = as.factor(batch))) +
  geom_point() +
  scale_color_manual(values = color_index) +
  ggtitle("PC1 vs PC2") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 3: PC1 Scores by Batch
p[[3]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC1, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC1", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC1 Scores; p=", round(pc1_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 4: PC2 Scores by Batch
p[[4]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC2, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC2", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC2 Scores; p=", round(pc2_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Assign colors to sample types
color_index2<- brewer.pal(8, "Dark2")[1:length(unique(pca_nc$Sample_type))]
# Create Plot 5: PC1 vs PC2 Colored by Sample Type
p[[5]] <- ggplot(pca_nc, aes(x = PC1, y = PC2, color = Sample_type)) +
    geom_point() +
    scale_color_manual(values = color_index2) +
    ggtitle("PC1 vs PC2 by Sample Type") +
    theme(plot.title = element_text(hjust = 0.5))
```

```{r,results='hide'}
##Modify zero
# Load necessary library
library(dplyr)

df_ft[, 3:152] <- apply(df_ft[, 3:152], 2, as.numeric)
summary(df_ft[, 3:152])


# calculate 10% & 25% quantile
threshold_10 <- apply(df_ft[, 3:152], 2, function(x) quantile(x, 0.10, na.rm = TRUE))
df_10 <- df_ft
for (i in 3:152) {
  df_10[, i] <- ifelse(df_10[, i] < threshold_10[i-2], 0, df_10[, i])
}

summary(df_10[, 3:152])
saveRDS(df_10, "data/df_10")

threshold_25 <- apply(df_ft[, 3:152], 2, function(x) quantile(x, 0.25, na.rm = TRUE))
df_25 <- df_ft
for (i in 3:152) {
  df_25[, i] <- ifelse(df_25[, i] < threshold_25[i-2], 0, df_25[, i])
}
saveRDS(df_25, "data/df_25")
```

```{r,results='hide'}
##10% dataset
df <- df_10
df <- df[, sapply(df, is.numeric)]
# PCA analysis
imputed_data <- imputePCA(df, ncp = 2)  
df_imputed <- imputed_data$completeObs 

pca_result <- prcomp(df_imputed, scale = TRUE)
pca_nc <- data.frame(pca_result$x)  
pca_nc <- pca_nc %>%
  mutate(sum_sample_int = rowSums(df_imputed))


pca_nc <- cbind(pca_nc, metadata)  # combine batch information

# ANOVA analysis
pc1_anova_nc <- summary(aov(PC1 ~ batch, data = pca_nc))
pc2_anova_nc <- summary(aov(PC2 ~ batch, data = pca_nc))
int_anova_nc <- summary(aov(sum_sample_int ~ batch, data = pca_nc))
# Initialize plot list
t <- list()

# Plot 1: Sum Intensity for All Samples

t[[1]] <- ggplot(pca_nc, aes(y = sum_sample_int, x = 1:length(sum_sample_int), fill = as.factor(batch))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = color_index) +
  labs(x = "Sample number", y = "Sum Intensity", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("Sum intensity for all samples; p=", round(int_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 2: PC1 vs PC2 Colored by Batch
t[[2]] <- ggplot(pca_nc, aes(x = PC1, y = PC2, color = as.factor(batch))) +
  geom_point() +
  scale_color_manual(values = color_index) +
  ggtitle("PC1 vs PC2") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 3: PC1 Scores by Batch
t[[3]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC1, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC1", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC1 Scores; p=", round(pc1_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 4: PC2 Scores by Batch
t[[4]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC2, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC2", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC2 Scores; p=", round(pc2_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r,results='hide'}
##25% zero
df <- df_10
df <- df[, sapply(df, is.numeric)]
# PCA analysis
imputed_data <- imputePCA(df, ncp = 2)  
df_imputed <- imputed_data$completeObs 

pca_result <- prcomp(df_imputed, scale = TRUE)
pca_nc <- data.frame(pca_result$x)  
pca_nc <- pca_nc %>%
  mutate(sum_sample_int = rowSums(df_imputed))


pca_nc <- cbind(pca_nc, metadata)  # combine batch information

# ANOVA analysis
pc1_anova_nc <- summary(aov(PC1 ~ batch, data = pca_nc))
pc2_anova_nc <- summary(aov(PC2 ~ batch, data = pca_nc))
int_anova_nc <- summary(aov(sum_sample_int ~ batch, data = pca_nc))
# Initialize plot list
s <- list()

# Plot 1: Sum Intensity for All Samples

s[[1]] <- ggplot(pca_nc, aes(y = sum_sample_int, x = 1:length(sum_sample_int), fill = as.factor(batch))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = color_index) +
  labs(x = "Sample number", y = "Sum Intensity", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("Sum intensity for all samples; p=", round(int_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 2: PC1 vs PC2 Colored by Batch
s[[2]] <- ggplot(pca_nc, aes(x = PC1, y = PC2, color = as.factor(batch))) +
  geom_point() +
  scale_color_manual(values = color_index) +
  ggtitle("PC1 vs PC2") +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 3: PC1 Scores by Batch
s[[3]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC1, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC1", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC1 Scores; p=", round(pc1_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))

# Plot 4: PC2 Scores by Batch
s[[4]] <- ggplot(pca_nc, aes(x = as.factor(batch), y = PC2, color = as.factor(batch))) +
  geom_jitter(position = position_dodge(0.8)) +
  scale_color_manual(values = color_index) +
  labs(x = "Batch Number", y = "PC2", size = 20) +
  guides(fill = guide_legend(title = "Batch Number")) +
  ggtitle(label = paste("PC2 Scores; p=", round(pc2_anova_nc[[1]]$`Pr(>F)`[1], 4), sep = " ")) +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#install
install.packages("plotly")
install.packages("shiny")
library(plotly)
library(shiny)
```


```{r last_chunk, echo = FALSE, results="show",message=FALSE}
##Interactive plot
# change ggplot to plotly
plist_interactive <- lapply(p, plotly::ggplotly)
tlist_interactive <- lapply(t, plotly::ggplotly)
slist_interactive <- lapply(s, plotly::ggplotly)

# UI
ui <- fluidPage(
  titlePanel("Interactive Plot Selector"),
  
  sidebarLayout(
    sidebarPanel(
      # panel
      selectInput("data_choice", "Select data group:",
                  choices = c("original data" = "p", 
                              "10% zero data" = "t", 
                              "25% zero data" = "s")),
      
      # buton
      actionButton("plot1", "sum intensity"),
      actionButton("plot2", "PC1 vs PC2"),
      actionButton("plot3", "PC1"),
      actionButton("plot4", "PC2")
    ),
    
    mainPanel(
      # display
      plotlyOutput("selected_plot")
    )
  )
)

# Server
server <- function(input, output, session) {
  
  # store via reactiveVal
  current_plot <- reactiveVal(1)
  
  # upload display
  selected_data <- reactive({
    if(input$data_choice == "p") {
      return(p)
    } else if(input$data_choice == "t") {
      return(t)
    } else {
      return(s)
    }
  })
  
  # upload chosen plot
  observeEvent(input$plot1, {
    current_plot(1)
  })
  
  observeEvent(input$plot2, {
    current_plot(2)
  })
  
  observeEvent(input$plot3, {
    current_plot(3)
  })
  
  observeEvent(input$plot4, {
    current_plot(4)
  })
  
  
  output$selected_plot <- renderPlotly({
    selected_data()[[current_plot()]]
  })
}

# run shinny
shinyApp(ui = ui, server = server)
library(htmlwidgets)
```
[View Source Code on GitHub](https://github.com/Yueran016/555_Dashboard)

##This presents three different data processing methods: Original Data, 10% Zero Data, and 25% Zero Data. Each data group is displayed through four types of plots: Sum Intensity, PC1 vs PC2, PC1 Distribution, and PC2 Distribution. These plots aim to help the reader understand the impact of different missing data treatments on data analysis and principal component analysis results, and further explore how missing data influences the overall data structure.